On Verification Patterns for Low-Level Systems via Modal Abstractions

### 2025-09-12 paper.tex
+

\subsubsection{Reference-Counting Memory Reclamation}
In a typical configuration, allocating, deallocating and referencing memory locations for client code 
leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory. DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
% To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
% \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, 
% from which a path is created for a stack pointer reference to the allocated memory. 
%\todo[inline,color=cyan]{
%Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
%Do we mean that it is ``often'' the case that it's a DAG?}

@@

To reclaim the allocated memory location that is no longer being accessed, many systems use a well-known method called \emph{reference-counting} in which 
each memory allocation includes not only its data but also a counter of how many pointers to that memory exist in the rest of the program.
Any time a new reference to the allocated memory is created, the  counter is incremented,
and every time an existing reference is destroyed (by overwriting or dropping) the counter is decremented.\footnote{This may be done
using library functions for a reference counting system, or the operations may be inserted by a compiler.}
These reference counts are a sound (but incomplete) way to approximate reachability: if a decrement of a reference count
decrements the count to 0, no other references to that memory exist in the program, and the memory can be reclaimed.

@@

\paragraph{Reference Counting} When dealing with reference-counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address. A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.

>>>>

\subsubsection{Reference Counting Memory Reclamation}
In a typical configuration, allocating, deallocating and referencing memory locations for client code 
leaves the memory layout \emph{ideally} in a form of DAG (directed acyclic graph) with a unique root, which resembles the first stack pointer referencing the allocated memory. DAG abstraction instead of recursive decent tree allows non-duplication on the reference counts for the identical memory objects.
% To ease picturing the memory layout after a set of requests from the client code (for allocation, referencing etc.), we can think of the unique roots of DAGs as 
% \emph{super} block addresses -- this time multiple ones, not only two -- in Figure \ref{fig:cowindex}, 
% from which a path is created for a stack pointer reference to the allocated memory. 
%\todo[inline,color=cyan]{
%Why is this a DAG? Normally programs can easily create cycles, so the DAG bit requires a little more explanation.
%Do we mean that it is ``often'' the case that it's a DAG?}

@@

To reclaim the allocated memory location that is no longer being accessed, many systems use a well-known method called \emph{reference counting} in which 
each memory allocation includes not only its data but also a counter of how many pointers to that memory exist in the rest of the program.
Any time a new reference to the allocated memory is created, the  counter is incremented,
and every time an existing reference is destroyed (by overwriting or dropping) the counter is decremented.\footnote{This may be done
using library functions for a reference counting system, or the operations may be inserted by a compiler.}
These reference counts are a sound (but incomplete) way to approximate reachability: if a decrement of a reference count
decrements the count to 0, no other references to that memory exist in the program, and the memory can be reclaimed.

@@

\paragraph{Reference Counting} When dealing with reference counting APIs, we may care to specify reachability of memory nodes ($\ell \mapsto v$) in a certain context defined by a shared root address. A shared memory address $\ell$ can be the root of the graph that can be a container of memory nodes ($\ell_1 \mapsto v$) that are reachable from the root $\ell$.
